Turbo fast no compromise gravitational wave parameter estimation

### 2022-12-08 src/tex/ms.tex
update figure

% Emerging tools from different group
While the standard tools can deliver trusted PE results, they are known to be
computationally intensive. The exact amount of time needed to analyze one event
depends on the length of the event. These community tools take about a week to
analyze shorter events such as binary black hole (BBH) mergers, where they can
take up to a month on a computer cluster to analyze longer events such as binary
neutron star (BNS) mergers. There are effort

@@

\kw{Dingo can't do BNS yet}

@@

where $\mathcal{L}(d|\theta)$ is the likelihood function, $\pi(\theta)$ is the
prior distribution, and $p(d)$ is the evidence. Since the evidence is a
normalization constant that does not depend on the source parameters, it is
often omitted if we are only interested in the posterior distribution. The prior
distribution is often chosen to be some simple distributions, such as uniformly
distributed in the component masses or a Gaussian distribution in the spins.
The most computationally intensive part of the PE is the evaluating the likelihood function.
Assuming the noise follows from a stationary Gaussian process, the log likelihood in the case of GW is given by

\begin{align}
    \log{\mathcal{L}(d|\theta)} = -\left<d-h(\mathbf{\theta})|d-h(\mathbf{\theta})\right>/2,
\label{eq: loglikelihood}
\end{align}
where $d$ is the data, $h(\mathbf{\theta})$ is the strain predicted by a waveform model with a specific set of source parameters.
It is computationally cheaper to compute the likelihood in the frequency domain, which can be defined as 
\begin{align}
    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
\label{eq: innerproduct}
\end{align}
where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD).

@@

The computational cost of evaluating a waveform model scales linearly with the
number of sample points either in time or frequency domain. This introduces
greater computational burden for longer-duration signals. To reduce the
computational cost, there are a number of methods to reduce the number of basis
points one would need to compute the likelihood faithfully \kw{cite}. We use
heterodyned likelihood \kw{cite} (also named relative binning in \kw{cite}) in
this work. We give a concise description of what we implement in our code here.
For more extensive discussion of the derivation of heterodyned likelihood, we
refer the reader the reference \kw{}.

\kw{Specify the idea of heterodyned likelihood here}.

@@

@@

@@

In this work, we present a PE pipeline for GW events that is efficient and flexible.
Our implemneted the following major features to achieve its performance:
\begin{enumerate}
\item Differential waveform models
\item Normalizing flow enhanced MCMC sampler
\item Heterodyned likelihood
\item Native support for accelerator
\item JIT compilation
\end{enumerate}

@@

As there are multiple group

@@

% Higher dimensional problems
While standard GW analysis goes up to 17 dimensions,

>>>>

% Emerging tools from different group
While the standard tools have passed many robustness tests, they are known to be
computationally intensive. The exact amount of time needed to analyze one event
depends on the length of the event. These community tools take about a week to
analyze shorter events such as binary black hole (BBH) mergers, where they can
take up to a month on a computer cluster to analyze longer events such as binary
neutron star (BNS) mergers. There are efforts from multiple groups developing
tools to speed up the PE process \kw{Cite}.

@@

In this work, we present a light-weighted, flexible and high performance
framework to estimate GW event parameters. Our framework implemneted the
following major features to achieve its performance:
\begin{enumerate}
\item Differential waveform models
\item Normalizing flow enhanced MCMC sampler
\item Heterodyned likelihood
\item Native support for accelerator
\item JIT compilation
\end{enumerate}
\kw{Dingo can't do BNS yet}

@@

where $\mathcal{L}(d|\theta)$ is the likelihood function, $\pi(\theta)$ is the
prior distribution, and $p(d)$ is the evidence. Since the evidence is a
normalization constant that does not depend on the source parameters, it is
often omitted if we are only interested in the posterior distribution. The prior
distribution is often chosen to be some simple distributions, such as uniformly
distributed in the component masses or a Gaussian distribution in the spins.
Assuming the noise follows from a stationary Gaussian process, the log likelihood in the case of GW is given by

\begin{align}
    \log{\mathcal{L}(d|\theta)} = -\left<d-h(\mathbf{\theta})|d-h(\mathbf{\theta})\right>/2,
\label{eq: loglikelihood}
\end{align}
where $d$ is the strain data, $h(\mathbf{\theta})$ is the strain predicted by a
waveform model with a specific set of source parameters. The right handside of
eq. \ref{eq: loglikelihood} can be evaluated in either time or frequency domain.
Since it is computationally cheaper to compute the likelihood in the frequency
domain, we choose to compute the likleihood throughout this work, which can be defined as 
\begin{align}
    \left<a|b\right> = 4 Re\int \frac{a^*(f)b(f)}{\mathcal{S}_n(f)} df,
\label{eq: innerproduct}
\end{align}
where $\mathcal{S}_n(f)$ is the noise power spectral density (PSD). To compute
the integral shown in eq. \ref{eq: innerproduct}, we need to evaluate a chosen
waveform model $h(\mathbf{\theta})$ at a number of frequnecy sample points.
Evaluting the likelihood function is often the most computationally intensive
part of the PE. The most accurate waveform model is numerical relativity (NR).
However, depending on the source parameters, generating one time series of
strain using can takes a day to half a year, which makes NR prohibitably
expensive for PE. To circumvent this problem, there are a number of waveform
"approximant" families, including the IMRPhenom family \kw{}, SEOB family \kw{},
and NR surrogate family\kw{cite}. For shorter event such as a $30-30\ M_{\odot}$
BBH, one waveform call could take $10\text{ms}$ to $\sim 1s$ \kw{verify number
in lalsuite}. For longer event such as a $1.4-1.4\ M_{\odot}$ BNS event, the
evaluation time could go up to \kw{Fill}.

@@

Since the computational cost of evaluating a waveform model scales linearly with
the number of sample points either in time or frequency domain, the
computational burden for longer-duration signals are often quite large. To
reduce the computational cost, there are a number of methods to reduce the
number of basis points one would need to compute the likelihood faithfully
\kw{cite}. We use heterodyned likelihood \kw{cite} (also named relative binning
in \kw{cite}) in this work. We give a concise description of what we implement
in our code here. For more extensive discussion of the derivation of heterodyned
likelihood, we refer the reader the reference \kw{}.

\kw{Specify the idea of heterodyned likelihood here}.
The idea behind heterodyned likelihood is the following:

@@

\begin{figure}
\includegraphics[width=0.99\linewidth]{static/GW150914.png}
\caption{
    hi
}
\label{fig:GW150914}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.99\linewidth]{static/GW170817.png}
\caption{
    hi
}
\label{fig:GW170817}
\end{figure}

\begin{figure}
\includegraphics[width=0.99\linewidth]{static/ppplot.png}
\caption{ \kw{I am pretty sure I am doing something wrong about the ppplot.
    Given injected q=1, there is no way the quantile is not 1, which is what
    bias the result seemingly.} }
\label{fig:ppplot}
\end{figure}

@@

% Brief summary
In this work, we present a PE pipeline for GW events that is efficient and flexible.

@@

@@

% Higher dimensional problems
While standard GW analysis goes up to 17 dimensions. non-standard GW PE problems
can have more parameters, which could potentially lead to more complicate
geometry in the target posterior. For example,
